ESRC SDAI Application - Understanding domestic abuse using Big Data

### 2019-07-10 JeS_fields.tex
Anna's really shitty summary

\begin{document}
\title{ESRC SDAI Application - Understanding domestic abuse using Big Data}
\date{}
\maketitle
\section{Objectives}
\textit{List the main objectives of the proposed research [up to 4000 chars]}
\section{Summary}

@@

\textit{Not sure what the right angle is, my current approach is to 1) demonstrate the extent of the problem 2) highlight the gaps in our current knowledge and 3) very briefly explain how are we contributing}

Domestic abuse is a complex, multifaceted phenomenon, that has a profound impact on the lives of those affected. Anyone can become a victim, regardless of socio-economic status. The official definition of domestic abuse in the UK aims a wide range of behaviours, including emotional, sexual, and physical abuse, threatening, intimidating, coercive and controlling behaviour \cite{govuk}.

Domestic abuse is increasingly recognised as a serious public health concern in many countries, including the UK. In the year ending 2018 March, an estimated 2 million adults in England and Wales (7.9\% of women and 4.2\% of men) have experienced some form of domestic abuse \cite{ONS}. Besides being more likely to experience any form of domestic abuse, compared to men, women are also significantly more likely to endure severe, often lethal violence at home: 41\% of femicides, and 0.6\% of homicides committed in the year ending 2018 March were domestic abuse-related \cite{homic}. The long-lasting effect of domestic abuse are not limited to the direct target of abuse, as it can have severe developmental impacts on children witnessing the abuse. Taking into account the emotional and physical consequences of the abuse suffered, the economic cost of domestic abuse in England and Wales between April, 2016 and March, 2017 was estimated to be \pounds 66 billion \cite{costs}.

Despite the vast economic costs associated with it, there is relatively little academic research focusing on domestic abuse in the context of the UK. This is most likely due to the scarcity of data sources relating to this topic. Owing to the fact that it occurs in private settings, domestic abuse is a vastly underreported crime: it is estimated that only 17\% of victims who experienced any form of domestic abuse between April 2017 and March 2018 reported it to the police \cite{ONS}, which means that studies exclusively drawing on police data to understand various aspects of domestic abuse (e.g. \citeNP{Bland2015}) have limited ability to derive useful insights about this phenomenon. Another methodological approach is to conduct interviews with survivors (e.g. \citeNP{callaghan}), which, while are tremendously helpful in illuminating the contexts in which domestic abuse can occur, are unfortunately limited in the generalisability of their findings.

In the UK, the most reliable source of data on domestic abuse is the Crime Survey for England and Wales (CSEW), an annual, cross-sectional representative victimisation survey across England and Wales, collecting data from 30,000--40,000 households every year \cite{OfficeforNationalStatistics2019}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular.  While the Office for National Statistics (ONS) regularly publishes findings from the CSEW, these are mostly exploratory and descriptive in nature. In spite of the rich information provided by this survey, a limited number of studies have utilised the CSEW (e.g., \citeNP{Khalifeh2013a}) to understand domestic abuse.

@@

>>>>

\begin{document}
\title{ESRC SDAI Application - Understanding domestic abuse using Big Data}
\date{}
\maketitle
\section{Objectives}
\textit{List the main objectives of the proposed research [up to 4000 chars]}

1) To conduct the first large-scale, quantitative exploration of the individual- and neighbourhood-level risk factors of domestic abuse victimisation. Rigorous validation of insights from existing, mostly quantitative investigations using an UK-wide dataset, providing a deeper understanding of the environments perpetuating domestic abuse. While we will only use UK data, these findings will likely to be relevant for researchers outside the UK.

2) To translate our research into impact by delivering applicable insights for the relevant agencies.  Police (improving recording practices, understanding domestic abuse reporting demand characteristics over time for a better allocation of resources), government (temporal/geographical planning of targeted campaigns, intervention programmes), shelters (temporal/geographical allocation of resources), academia (publications)

\section{Summary}

@@

\textit{I'm really bad at this. I don't really understand what should go in here}

Who are the people most vulnerable to becoming victims of domestic abuse? To what extent do neighbourhood characteristics predict the prevalence of domestic abuse? How can we prevent serious harm and what affects victims' decision to finally report the abuse? What is the quantifiable, long-term impact of domestic abuse victimisation, or witnessing domestic abuse as a child on socio-economic outcomes in later life? How do changes in alcohol consumption and financial stability affect the prevalence of domestic abuse?

Addressing these questions are fundamental for a deeper understanding of the constellation of individual and environmental factors perpetuating domestic abuse, and the long-term consequences it can have on the lives of those affected. Designing effective policy interventions to protect victims can only be based on a sound understanding of the type of environments where domestic abuse is likely to occur. In this research programme, we combine various data sources to provide a comprehensive answer to these questions in the context of the UK.

Domestic abuse is increasingly recognised as a serious public health concern in many countries, including the UK. It is estimated that in the year ending 2018 March, 2 million adults in England and Wales (7.9\% of women and 4.2\% of men) have experienced some form of domestic abuse \cite{ONS}. Escalation of domestic abuse can result in homicide, and the risk of lethal violence is especially high in the case of female victims: 41\% of femicides, and 0.6\% of homicides committed in the year ending 2018 March were domestic abuse-related in England and Wales \cite{homic}. The long-lasting effect of domestic abuse are not limited to the direct target of abuse, as it can have severe developmental impacts on children witnessing the abuse. Taking into account the emotional and physical consequences of the abuse suffered, the economic cost of domestic abuse in England and Wales between April, 2016 and March, 2017 was estimated to be \pounds 66 billion \cite{costs}.

Despite the vast economic costs associated with it, there is relatively little academic research focusing on domestic abuse in the context of the UK. Studies using police-recorded domestic abuse data have limited ability to derive useful insights about this phenomenon due to the vast levels of underreporting. In contrast, studies using qualitative methods have limited ability to derive useful insights about this phenomenon.

%Domestic abuse is a complex, multifaceted phenomenon, that has a profound impact on the lives of those affected. It is increasingly recognised as a serious public health concern in many countries, including the UK. In the year ending 2018 March, an estimated 2 million adults in England and Wales (7.9\% of women and 4.2\% of men) have experienced some form of domestic abuse \cite{ONS}. Besides being more likely to experience any form of domestic abuse, compared to men, women are also significantly more likely to endure severe, often lethal violence at home: 41\% of femicides, and 0.6\% of homicides committed in the year ending 2018 March were domestic abuse-related \cite{homic}. The long-lasting effect of domestic abuse are not limited to the direct target of abuse, as it can have severe developmental impacts on children witnessing the abuse. Taking into account the emotional and physical consequences of the abuse suffered, the economic cost of domestic abuse in England and Wales between April, 2016 and March, 2017 was estimated to be \pounds 66 billion \cite{costs}.

%Despite the vast economic costs associated with it, there is relatively little academic research focusing on domestic abuse in the context of the UK. This is most likely due to the scarcity of data sources relating to this topic. Owing to the fact that it occurs in private settings, domestic abuse is a vastly underreported crime: it is estimated that only 17\% of victims who experienced any form of domestic abuse between April 2017 and March 2018 reported it to the police \cite{ONS}, which means that studies exclusively drawing on police data to understand various aspects of domestic abuse (e.g. \citeNP{Bland2015}) have limited ability to derive useful insights about this phenomenon. Another methodological approach is to conduct interviews with survivors (e.g. \citeNP{callaghan}), which, while are tremendously helpful in illuminating the contexts in which domestic abuse can occur, c.

In the UK, the most reliable source of data on domestic abuse is the Crime Survey for England and Wales (CSEW), an annual, cross-sectional representative victimisation survey across England and Wales, collecting data from 30,000--40,000 households every year \cite{OfficeforNationalStatistics2019}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular.  While the Office for National Statistics (ONS) regularly publishes findings from the CSEW, these are mostly exploratory and descriptive in nature. In spite of the rich information provided by this survey, only a limited number of studies have utilised the CSEW (e.g., \citeNP{Khalifeh2013a}) to understand domestic abuse.

@@

para on implications
